---
title: "Assignment4"
author: "Rona Hu"
date: "`r format(Sys.time(), '%B %Y')`"
output:
  html_document:
    theme: spacelab
    highlight: tango
    toc: yes
    number_sections: no
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: yes
    code_folding: show
    self_contained: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
  word_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include = FALSE, tidy = TRUE}
knitr::opts_chunk$set(echo = TRUE,cache = TRUE, 
                      autodep = TRUE, cache.comments = FALSE,
                      message = FALSE, warning = FALSE,
                      fig.width = 8, fig.height = 8)
```

##### The Github link:
https://github.com/RonaGitHub/Hu1-a1.git


```{r}
# Necessary libraries for this project
# install.packages("bigrquery")
library(bigrquery)
library(tidyverse)
library(DBI)
library(dbplyr)
```

In this notebook we will use Google BigQuery, "Google's fully managed, petabyte scale, low cost analytics data warehouse". Some instruction on how to connect to Google BigQuery can be found here: https://db.rstudio.com/databases/big-query/.

You will need to set up a Google account with a project to be able to use this service. We will be using a public dataset that comes with 1 TB/mo of free processing on Google BigQuery. As long as you do not repeat the work in this notebook constantly, you should be fine with just the free tier.

Go to https://console.cloud.google.com and make sure you are logged in a non-university Google account. This may not work on a university G Suite account because of restrictions on those accounts. Create a new project by navigating to the dropdown menu at the top (it might say "Select a project") and selecting "New Project" in the window that pops up. Name it something useful.

After you have initialized a project, paste your project ID into the following chunk.

```{r project}
project <- "survmeth727"
```

We will connect to a public database, the Chicago crime database, which has data on crime in Chicago.

```{r}
## Solve the authentication problem
#library(gargle)

# Clear the old token by setting it to NULL
#options("gargle_oauth_cache" = ".secrets",
#        "gargle_oauth_email" = "(hide)")

# Define the required scopes
#scopes <- c(
#  "https://www.googleapis.com/auth/bigquery",
#  "https://www.googleapis.com/auth/cloud-platform"
#)

# Re-authenticate with the new account and scopes
# bq_auth(path = NULL, email = "(hide)", scopes = scopes)

# Follow the prompts in your web browser to authenticate with the new account


con <- dbConnect(
  bigrquery::bigquery(),
  project = "bigquery-public-data",
  dataset = "chicago_crime",
  billing = project
)
con
```

We can look at the available tables in this database using dbListTables.

Note: When you run this code, you will be sent to a browser and have to give Google permissions to Tidyverse API Packages. Make sure you select all to give access or else your code will not run.

```{r}
dbListTables(con)
```

```{r, message=FALSE}
dbListTables(con) #see names of tables in database
dbListFields(con, "crime") #see column names
```


Information on the `crime` table can be found here:

https://cloud.google.com/bigquery/public-data/chicago-crime-data

#### Using SQL Code

Write a first query that counts the number of rows of the `crime` table in the year 2016. Use code chunks with {sql connection = con} in order to write SQL code within the document.

```{sql connection= con, message = FALSE}
SELECT count(*)
FROM crime
WHERE year = 2016
LIMIT 10;
```


Next, count the number of arrests grouped by primary_type in 2016. Note that is a somewhat similar task as above, with some adjustments on which rows should be considered. Sort the results, i.e. list the number of arrests in a descending order.

```{sql connection= con, message = FALSE}
SELECT count(arrest), primary_type
FROM crime
WHERE year = 2016 AND arrest = TRUE
GROUP BY primary_type
ORDER BY count(arrest) DESC
LIMIT 10;
```

We can also use the date for grouping. Count the number of arrests grouped by hour of the day in 2016. You can extract the latter information from date via EXTRACT(HOUR FROM date). Which time of the day is associated with the most arrests?

```{sql connection= con, message = FALSE}
SELECT count(arrest), EXTRACT(HOUR FROM date)
FROM crime
WHERE year = 2016 AND arrest = TRUE
GROUP BY EXTRACT(HOUR FROM date)
LIMIT 15;
```

*Answer*: The time 12 of the day has the most arrests.

Focus only on HOMICIDE and count the number of arrests for this incident type, grouped by year. List the results in descending order.

```{sql connection= con, message = FALSE}
SELECT COUNT(arrest) AS arrest_count, year
FROM crime
WHERE primary_type = 'HOMICIDE' AND arrest = TRUE
GROUP BY year
ORDER BY arrest_count DESC
LIMIT 10;
```

Find out which districts have the highest numbers of arrests in 2015 and 2016. That is, count the number of arrests in 2015 and 2016, grouped by year and district. List the results in descending order.

```{sql connection= con, message = FALSE}
SELECT COUNT(arrest) AS arrest_count, year, district
FROM crime
WHERE (year = 2015 OR year = 2016) AND arrest = TRUE
GROUP BY year, district
ORDER BY arrest_count DESC
LIMIT 10;
```

#### Using DBI Package

Lets switch to writing queries from within R via the DBI package. Create a query object that counts the number of arrests grouped by primary_type of district 11 in year 2016. The results should be displayed in descending order.

```{r}
query <- "SELECT EXTRACT(YEAR FROM date) as Year, primary_type, COUNT(arrest) as count_of_arrests FROM crime
          WHERE district = 11 AND Year = 2016
          GROUP BY primary_type, Year
          ORDER BY count_of_arrests DESC"

# Execute the query.
result <- dbGetQuery(con, query)

result
```

Try to write the very same query, now using the dbplyr package. For this, you need to first map the crime table to a tibble object in R.

```{r}
# Assign data table to object
crime <- tbl(con, "crime")

# Check to see if object is tibble
str(crime)
class(crime)
```

Again, count the number of arrests grouped by primary_type of district 11 in year 2016, now using dplyr syntax.

```{r}
crime %>%
  select(arrest, primary_type, district, year) %>%
  filter(district == 11 & year == 2016 & arrest == TRUE) %>%
  group_by(primary_type) %>% 
  count()
```

Count the number of arrests grouped by primary_type and year, still only for district 11. Arrange the result by year.
Assign the results of the query above to a local R object.
Confirm that you pulled the data to the local environment by displaying the first ten rows of the saved data set.

```{r, message = FALSE}
crime_tbl <- tbl(con, "crime")

glimpse(crime_tbl)

crime_tbl <- crime_tbl %>%
  mutate(year = year(date))

query2 <- crime_tbl %>%
  filter(district == 11) %>%
  group_by(primary_type, year) %>%
  summarize(count_of_arrests = n()) %>%
  arrange(desc(year))
# 
arrests_by_year <- query2 %>% collect()

arrests_by_year
```

Close the connection.

```{r}
dbDisconnect(con)
```

